### 2014-01-20 paper/manuscript.tex
Third version of the paper, with figures.

%% \newcommand{\hidefigures}{\relax}

@@

\ \\
{\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Ontario, Canada.}\\
%

@@

Noise and heterogeneity are both known to benefit neural coding. Stochastic resonance describes how noise, in the form of random fluctuations in a neuron's membrane voltage, can improve neural representations of an input signal. Neuronal heterogeneity refers to variation in any one of a number of neuron parameters, and is also known to increase the information content of a population. We explore the interaction between noise and heterogeneity, and find that their benefits to neural coding are not independent. Specifically, a neuronal population better represents an input signal when either noise or heterogeneity is added, but adding both does not always improve representation further. To explain this phenomenon, we propose noise and heterogeneity both operate using two shared mechanisms: 1) temporally desynchronizing the firing of neurons in the population, and 2) linearizing the response of a population to a stimulus. We first characterize the effects of noise and heterogeneity on the information content of populations of either leaky integrate-and-fire or FitzHugh-Nagumo neurons. We then examine how the mechanisms of desynchronization and linearization produce these effects, and find that they work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization). Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows neuron to encode \emph{complementary} aspects of the input signal, thereby increasing information content. The simulations detailed in this paper highlight the importance of heterogeneity and noise in population coding, demonstrate their complex interactions in terms of the information content of neurons, and explain these effects in terms of underlying mechanisms.

@@

The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}. In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \emph{i.e.,} input signals which are entirely below the system's threshold. In contrast, a noisy input signal (\emph{i.e,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal. Too much noise, though, causes the input signal to become lost in random fluctuations. A moderate amount of noise---not too much, and not too little---is needed to optimize the information about the input signal contained in the system \citep{Wiesenfeld1994,Longtin1998}.

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise increases information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

The fact that neurons are heterogeneous---with variation across a wide range of parameters---has largely been ignored in the stochastic resonance literature. It is well known that neurons vary widely across the nervous system. Even neurons in the same brain area and of the same qualitative type vary across numerous parameters. Though this diversity in neural systems was documented by some of the first researchers in the field, it is only recently that neuroscientists have begun to illuminate the functional role of heterogeneity. Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal, and the range of functions computable by the population. More recent work showed that heterogeneity allows neurons to use combinatorial coding schemes \citep{Osborne2008} and in general increase the information encoded by a population \citep{Shamir2006,Chelaru2008,Padmanabhan2010,Ecker2011}. In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}. Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.

@@

We explore this hypothesis in the context of two neuron models: the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, and the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}. We drive the population of FHN or LIF neurons with a common aperiodic random signal $s(t)$, and vary the levels of internal noise and heterogeneity in the population. Finally, we decode an output signal from the population by summing and filtering the spikes from individual neurons, and compare the similarity of the decoded output signal with the input signal using mutual information. Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \textsc{\nameref{scn:methods}}). Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width.

We report how noise and heterogeneity both act to increase information content in these populations, but in a complex, non-independent manner. They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus, and desynchronizing neuronal firing---that we examine in detail.

@@

We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons. Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average). Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron. For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of band-limited aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$. This signal has equal power below 5 Hz and zero power above 5 Hz, where the threshold of 5 Hz was chosen because it is significantly slower (approximately $\frac{1}{5}$) of the maximum firing rate of the FHN and LIF neurons (25--30 Hz). The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input. The FHN or LIF neurons were simulated using their respective differential equations, and the output was decoded using a simple summing neuron that summed and low-pass filtered the outputs of the individual neurons to obtain the population output (outputs from ``off'' neurons were multiplied by -1 before summing). Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated. The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels. See \textsc{\nameref{scn:methods}} for the full details of the experimental setup.

Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of LIF neurons. At low noise levels, we observe a resonance effect in the heterogeneity parameter---there is an optimal level of heterogeneity that maximizes the information content of the population. Furthermore, this optimal level of heterogeneity is around one to two times the root mean square amplitude of the input signal. This indicates that optimal heterogeneity corresponds to having neuron firing thresholds that span the full range of input signal values. Too little heterogeneity means that some input signal components cannot be uniquely encoded by the population, since these values will be either well below or well above the firing thresholds of all neurons in the population, and all neurons will fire approximately constantly for these values. Too much heterogeneity results in some neurons that have firing thresholds well below or well above all signal values, causing them to either fire tonically independent of the input signal value, or to fire for no part of the input signal, respectively. This type of resonance in heterogeneity was recently observed by \cite{Mejias2012}, both theoretically and in numerical experiments.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{../figures/bw/infohetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information content of a population of LIF neurons, and exhibits a resonance effect.}
    Adding heterogeneity to a population of LIF neurons increases their ability to represent an input signal, but only for lower levels of noise. At higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
  }
  \label{fig:infohetero}
\end{figure}

At higher levels of noise, heterogeneity offers no benefit to LIF neuron populations (Figure~\ref{fig:infohetero}). To examine the interactions between noise and heterogeneity further, we conducted the same experiment across a wide range of noise and heterogeneity values. Figure~\ref{fig:infocontour} illustrates the effects of these interactions on the information content of a population of LIF or FHN neurons. In LIF neurons, the population is best able to represent the input signal with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$) and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$). Given this optimal level of heterogeneity, adding noise only serves to decrease the information content of the system. But if no heterogeneity is present in the system, then adding the right amount of noise ($\sigma_\eta \approx 1.0 \times 10^{-2}$) optimizes information content (Figure~\ref{fig:infonoise}). This shows that the effects of noise and heterogeneity are not independent.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity show nonlinear interactions in their benefits to information representation.} We measured the mutual information (contours every 0.15 bits) across many levels of noise and heterogeneity. In LIF neurons, both noise and heterogeneity show independent resonance effects: given that only one of noise or heterogeneity is present, there is a non-zero and finite level of noise or heterogeneity that optimizes information content. Setting both noise and heterogeneity to their optimal levels fails to further improve the representational ability of the population. Both parameters use the same underlying mechanisms of desynchronization and linearization, so once the population has been desynchronized and linearized by one parameter, adding the other cannot further improve information content. Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}), so information content in these neurons is maximized for a particular level of noise.
  }
  \label{fig:infocontour}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
  \fi
  \caption{
    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} In both LIF (left) and FHN (right) populations, noise exhibits a stochastic resonance effect, where adding noise to the homogeneous population (black line) improves the population's information content. This stochastic resonance effect disappears in heterogeneous populations of LIF neurons, but still occurs in heterogeneous populations of FHN neurons.
  }
  \label{fig:infonoise}
\end{figure}

We propose that noise and heterogeneity both improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population, and 2) linearizing the response of the population to an input signal. These shared mechanisms result in the nonlinear interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation. In the following two sections, we present additional experimental results to investigate and quantify the mechanisms of desynchronization and linearization.

We also note that there is a difference between FHN and LIF neurons in terms of noise and heterogeneity, and their effects on information content. Heterogeneity is much less effective at increasing information content in FHN neurons than LIF neurons (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}). This is explained by the inability of heterogeneity to fully desynchronize firing in FHN neurons (see \textsc{\nameref{scn:desync}}), which is in turn due to the nature of the FHN tuning curves. We revisit this result in the \textsc{\nameref{scn:discussion}}.

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
  \fi
  \caption{
    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 64 LIF neurons. With low noise and heterogeneity, neuronal firing is highly synchronized. Adding a moderate amount of noise or heterogeneity desynchronizes this firing. Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized. High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal. High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability. Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high). Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
  }
  \label{fig:syncraster}
\end{figure}

This previous example is, of course, artificial. It requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions. In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time. This is not the case, because signals below the firing thresholds of a group of neurons have a synchronizing effect on the group. By stopping neuronal firing, these signals push all the neurons in the group towards their stable equilibrium points. When the signal returns to an amplitude above the firing thresholds of neurons in the group, all neurons will begin firing at approximately the same time, and will be once again in phase. Signals with higher frequency components are able to cross larger ranges of firing thresholds in quick succession, and have a stronger synchronizing effect than purely low frequency signals. Thus, more negative (positive) portions of the input signal repeatedly act to synchronize ``on'' (``off'') neurons, and groups of neurons must be able to desynchronize quickly after these synchronization events in order to remain out of phase with one another.

@@

High frequency input signals may cross the thresholds of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously. LIF neurons are still able to desynchronize quickly after such a stimulus, since the LIF tuning curves used have different firing rates for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}). Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize. FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate. This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (Figure~\ref{fig:phase}B).

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/tuningnoisy.eps}
  \fi
  \caption{
    \textbf{Noise linearizes neuron tuning curves.} We probed LIF and FHN neurons using constant inputs at various voltages, using varying levels of noise, and measured the output firing rate. Noise has the effect of making the tuning curves more linear, specifically eliminating the sharp threshold in both LIF and FHN neurons. The main mechanism is that of subthreshold stochastic resonance: noise allows subthreshold inputs to cause spikes by occasionally pushing the membrane voltage above the threshold. This occurs in both both neuron models, allowing them to represent subthreshold signal components. In FHN neurons, small levels of noise can also cause firing rates for small suprathreshold signals to drop slightly by briefly pushing the neuron away from its firing limit cycle in state space, though interestingly this does not happen for larger noise levels. Large levels of noise in FHN neurons cause large membrane voltage fluctuations, resulting in jumps from one point on the limit cycle to another point much farther ahead, causing elevated firing rates. In LIF neurons, noise added to small suprathreshold signals causes the neuron to fire slightly faster, because the noise can cause the rising membrane voltage to prematurely cross the threshold, causing the spike to occasionally occur sooner than it would in the noiseless case.
  }
  \label{fig:tuning}
\end{figure}

>>>>

\graphicspath{{../figures/bw/}}
%% \newcommand{\hidefigures}{\relax}

@@

\ \\
{\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Canada.}\\
{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Canada.}\\
%

@@

Noise and heterogeneity are both known to benefit neural coding.
Stochastic resonance describes how noise,
in the form of random fluctuations in a neuron's membrane voltage,
can improve neural representations of an input signal.
Neuronal heterogeneity refers to variation in any one of a number of neuron parameters,
and is also known to increase the information content of a population.
We explore the interaction between noise and heterogeneity
and find that their benefits to neural coding are not independent.
Specifically, a neuronal population better represents an input signal
when either noise or heterogeneity is added,
but adding both does not always improve representation further.
To explain this phenomenon, we propose
that noise and heterogeneity both operate using two shared mechanisms:
1) temporally desynchronizing the firing of neurons in the population,
and 2) linearizing the response of a population to a stimulus.
We first characterize the effects of noise and heterogeneity
on the information content of populations of
either leaky integrate-and-fire or FitzHugh-Nagumo neurons.
We then examine how the mechanisms
of desynchronization and linearization produce these effects,
and find that they work to distribute information equally across all neurons in the population,
both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization).
Without noise or heterogeneity,
all neurons in the population encode the same aspects of the input signal;
adding noise or heterogeneity allows neurons
to encode \emph{complementary} aspects of the input signal,
thereby increasing information content.
The simulations detailed in this paper
highlight the importance of heterogeneity and noise in population coding,
demonstrate their complex interactions in terms of the information content of neurons,
and explain these effects in terms of underlying mechanisms.

@@

The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}. In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals (\emph{i.e.,} input signals which are entirely below the system's threshold). In contrast, a noisy input signal (\emph{i.e.,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allow the system to transmit more information about the input signal. Too much noise, however, causes the input signal to become lost in random fluctuations. A moderate amount of noise---not too much, and not too little---is needed to optimize the information about the input signal contained in the system \citep{Wiesenfeld1994,Longtin1998}.

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units and found that noise increases information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

The fact that neurons are heterogeneous---with variation across a wide range of parameters---has largely been ignored in the stochastic resonance literature. It is well known that neurons vary widely across the nervous system. Even neurons in the same brain area and of the same qualitative type vary across numerous parameters. Though this diversity in neural systems was documented by some of the first researchers in the field, it is only recently that neuroscientists have begun to illuminate the functional role of heterogeneity. Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal and the range of functions computable by the population. More recent work showed that heterogeneity allows neurons to use combinatorial coding schemes \citep{Osborne2008} and in general increase the information encoded by a population \citep{Shamir2006,Chelaru2008,Padmanabhan2010,Ecker2011}. In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but only in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000} or for very large levels of noise \citep{McDonnell2006}. Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.

@@

We explore this hypothesis in the context of two neuron models:
the FitzHugh-Nagumo (FHN) neuron model,
a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001},
and the leaky integrate-and-fire (LIF) neuron model,
a typical model of cortical neurons \citep{Koch1999}.
We drive the population of FHN or LIF neurons
with a common aperiodic random signal $s(t)$,
and vary the levels of internal noise and heterogeneity in the population.
Finally, we decode an output signal from the population
by summing and filtering the spikes from individual neurons
and compare the similarity of the decoded output signal
with the input signal using mutual information.
Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \textsc{\nameref{scn:methods}}). Heterogeneity is added by randomly selecting
the bias voltages $b_i$ from a uniform interval of varying width.
This type of heterogeneity has been used in previous studies \citep[e.g.,][]{Brody2003},
and captures the inherent variation in excitability among neurons in vivo \citep{Mejias2012} due to variations in neuron physiology or differences in the background activity of afferent neurons.

We report how noise and heterogeneity both act to increase information content in these populations, but in a complex, non-independent manner. They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus and desynchronizing neuronal firing---that we examine in detail.

@@

We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons. Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average). Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron. For the remainder of this paper, we will use the term ``noise'' to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 100 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$.
The aperiodic signal was created by convolving a Gaussian white noise signal with an $\alpha$ function ($\alpha(t) = (t / \tau) e^{-t / \tau}$), with the correlation time $\tau = 20$ ms \citep{Mainen1995}.
The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input. The FHN or LIF neurons were simulated using their respective differential equations, and the output was decoded using a simple summing neuron that summed and low-pass filtered the outputs of the individual neurons to obtain the population output (outputs from ``off'' neurons were multiplied by -1 before summing). Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated. The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 100 trials at those levels.
We also performed experiments for different sizes of neuronal population, and found that population size does not significantly affect the results presented here. Using smaller populations simply lowers the overall information content of the system slightly and makes single-trial results more variable because there are fewer neurons to average across.
The experimental setup is summarized in Figure~\ref{fig:exp}; see \textsc{\nameref{scn:methods}} for further details.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{experiment.eps}
  \fi
  \caption{
    \textbf{The experimental setup.} In our experiment, neurons are fed a common aperiodic input. Each neuron also has an independent white noise process and a randomly chosen bias that are added to the input. Varying the amplitude of the noise and the width of the interval from which the biases are randomly drawn allows us to control the amount of noise and heterogeneity in the experiment, respectively. Additionally, some ``off'' neurons invert their input (not shown). The output of each FHN or LIF neuron model is summed and filtered to produce the population output; this is compared with the input signal using mutual information as a metric.
  }
  \label{fig:exp}
\end{figure}

Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of neurons. At low noise levels, we observe a resonance effect in the heterogeneity parameter---there is an optimal level of heterogeneity that maximizes the information content of the population. Furthermore, this optimal level of heterogeneity is around one to two times the root-mean-square amplitude of the input signal. This indicates that optimal heterogeneity corresponds to having neuron firing thresholds that span the full range of input signal values. Too little heterogeneity means that some input signal components cannot be uniquely encoded by the population, since these values will be either well below or well above the firing thresholds of all neurons in the population, and all neurons will fire approximately constantly for these values. Too much heterogeneity results in some neurons that have firing thresholds well below or well above all signal values, causing them to either fire tonically independent of the input signal value, or to fire for no part of the input signal, respectively. This type of resonance in heterogeneity was recently observed by \cite{Mejias2012}, both theoretically and in numerical experiments.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infohetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information content of a population of neurons, and exhibits a resonance effect.}
    Adding some heterogeneity (x-axis) to a population of LIF or FHN neurons increases their ability to represent an input signal (y-axis), while too much heterogeneity decreases representational ability. This effect is only observed at lower noise levels; at higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
  }
  \label{fig:infohetero}
\end{figure}

Higher levels of noise reduce or completely eliminate the benefits offered by heterogeneity (Figure~\ref{fig:infohetero}). To examine the interactions between noise and heterogeneity further, we conducted the same experiment across a wide range of noise and heterogeneity values. Figure~\ref{fig:infocontour} illustrates the effects of these interactions on the information content of a population of LIF or FHN neurons. We measured both the absolute information content (in bits) and the information content per spike, or information rate (in bits per spike).

LIF neurons are best able to represent the input signal
with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$)
and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$).
This was true when measuring both the absolute information content and the information per spike.
FHN neurons, on the other hand, required both heterogeneity and noise
to achieve maximum information content,
because heterogeneity alone is not able to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}).
However, heterogeneity alone was able to maximize the information rate of the system.

We explain this difference between the information content and information rate of FHN neurons
with the fact that noise increases the firing rates of neurons.
While noise offers some benefit to the information content of a population of FHN neurons
(given an optimal level of heterogeneity),
this effect disappears when examining information per spike,
since noise also increases the number of spikes.
In Figure~\ref{fig:infonoise}, we examine more closely the effects of noise on information content and rate in the absence of heterogeneity. Noise shows a clear resonance effect when measuring the absolute information content of the population, but this resonance effect is reduced when measuring the information per spike. We conclude that noise operates partially by increasing the firing rates of neurons, and indeed this effect can be observed when measuring the tuning curves of neurons under noise (see Figure~\ref{fig:tuning}).

The information landscapes of Figure~\ref{fig:infocontour} demonstrate that the effects of noise and heterogeneity are not independent. This is most apparent when examining the absolute information content of LIF populations (top left), and the information rate of FHN populations (bottom right).
For example, LIF neurons show a resonance in heterogeneity in the absence of noise,
but they also show a resonance in noise in the absence of heterogeneity (see also Figure~\ref{fig:infonoise}).
However, if we choose the optimal levels of noise and heterogeneity in the absence of the other factor ($\sigma_\eta \approx 10^{-2}$ and $b_r \approx 2 \times 10^{-1}$), this combined level of noise and heterogeneity does not further increase the information content of the system. This indicates that the effects of noise and heterogeneity are not independent: the optimal level of noise depends on the level of heterogeneity, and vice versa.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
  \fi
  \caption{
    \textbf{The benefits of noise and heterogeneity to information representation are not independent.}
We measured the information content (top, in bits) and information rate (bottom, in bits per spike),
across many levels of noise and heterogeneity.
In LIF neurons, both noise and heterogeneity show independent resonance effects:
given that only one of noise or heterogeneity is present,
there is a non-zero and finite level of noise or heterogeneity that optimizes information content.
Setting both noise and heterogeneity to their (independent) optimal levels
fails to further improve the representational ability of the population,
showing that these effects are not independent.
We propose that both parameters use
the same underlying mechanisms of desynchronization and linearization,
so once the population has been desynchronized and linearized by one parameter,
adding the other cannot further improve information content.
Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}),
so information content in these neurons is maximized for a specific combination of heterogeneity and noise. However, this effect is not seen when examining the information rate,
indicating that noise benefits neural systems partly through increasing firing rates.
  }
  \label{fig:infocontour}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
  \fi
  \caption{
    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} A) In homogeneous populations, noise exhibits a resonance effect on information content. B) The resonance effect of noise is less pronounced when measuring the information rate, indicating that noise operates partially by increasing firing rates.
  }
  \label{fig:infonoise}
\end{figure}

We propose that both noise and heterogeneity improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population and 2) linearizing the response of the population to an input signal. These shared mechanisms result in the non-independent interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation. In the following two sections, we present additional experimental results to investigate and quantify the mechanisms of desynchronization and linearization.

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
  \fi
  \caption{
    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 32 LIF neurons. With low noise and heterogeneity, neuronal firing is highly synchronized. Adding a moderate amount of noise or heterogeneity desynchronizes this firing. Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized. High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal. High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability. Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high). Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
  }
  \label{fig:syncraster}
\end{figure}

This previous example is, of course, artificial. It requires that all neurons are \emph{exactly} the same and start with \emph{exactly} the same initial conditions. In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time. This is not the case, because signals below the firing thresholds of a group of neurons have a synchronizing effect on the group. By stopping neuronal firing, these signals push all the neurons in the group towards their stable equilibrium points. When the signal returns to an amplitude above the firing thresholds of neurons in the group, all neurons will begin firing at approximately the same time, and will be once again in phase. Signals with higher frequency components are able to cross larger ranges of firing thresholds in quick succession, and have a stronger synchronizing effect than purely low frequency signals. Thus, more negative (positive) portions of the input signal repeatedly act to synchronize ``on'' (``off'') neurons, and groups of neurons must be able to desynchronize quickly after these synchronization events in order to remain out of phase with one another.

@@

High frequency input signals may cross the thresholds of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously. LIF neurons are still able to desynchronize quickly after such a stimulus, since the LIF tuning curves used have different firing rates for different sizes of suprathreshold signal (Figure~\ref{fig:tuning}). Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize. FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (Figure~\ref{fig:tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate. This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (Figure~\ref{fig:phase}A), whereas the desynchronizing effect of heterogeneity in FHN neurons is more modest (Figure~\ref{fig:phase}B).

This difference in the ability of heterogeneity to desynchronize LIF versus FHN neurons
explains why LIF neurons achieve maximum information content with only heterogeneity,
whereas FHN neurons require both heterogeneity and noise (Figure~\ref{fig:infocontour}).
Since heterogeneity does not desynchronize FHN neurons as well as noise does,
adding noise to a heterogeneous population of FHN neurons further desynchronizes them
and increases information content.

@@

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/tuningnoisy.eps}
  \fi
  \caption{
    \textbf{Noise linearizes neuron tuning curves.}
    We measured the tuning (stimulus-response) curves of LIF and FHN neurons under various levels of noise.
    We observe that noise linearizes the neuron tuning curves,
    Specifically, noise eliminates the sharp threshold in the tuning curve by
    occasionally pushing the neuron above its firing threshold
    for stimuli that are below the threshold.
    Noise also increases neuron firing rates across some or all stimuli.
  }
  \label{fig:tuning}
\end{figure}

Neuron heterogeneity has a linearizing effect similar to adding noise to a population, but unlike the noisy case, where it made sense to talk about noise linearizing an individual neuron's tuning curve, here we will discuss how heterogeneity linearizes the \emph{population's} tuning curve. The population's tuning curve is the set of points connecting the input signal strength to the decoded output. Since our model uses a simple summing decoder, the population tuning curve is equal to the sum of the individual neuron tuning curves. For a population of neurons with uniformly varied random firing thresholds, the population tuning curve is much more linear than a single neuron's tuning curve, whereas in the homogeneous case the population's tuning curve is the same as the neurons' tuning curve (Figure~\ref{fig:tuninghetero}). The linearizing effect of heterogeneity thereby allows for better representation of an input signal. This result is related to the original idea from \cite{Hubel1962} that variation in neural receptive fields provides a better basis for representing an input signal, and reappears in the recent heterogeneity literature \citep{Angelo2012}.
